---
title: "CS1950_Week03_PiB"
author: "Modified by Hunsi" 
date: "1/29/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages

```{r, load_tidyverse_pkg}
library(tidyverse)
```

## Read data

```{r, read_csv_file}
file_name <- ("https://raw.githubusercontent.com/jyurko/cs_1950_brain_proj/main/KLU_APC2_Master_2021_07_07.csv?token=GHSAT0AAAAAABQQ7BMCOBSTBU3XSIKIUTKEYQGUGLQ")
df <- readr::read_csv(file_name,col_names = TRUE)
```

## Check unique identifiers

The number of rows associated with each `Vault_Scan_ID`.  

```{r, count_scan_ids_a}
df %>% 
  count(Vault_Scan_ID)
```

The count for the number of rows associated with each `Vault_Scan_ID`.  

```{r, count_scan_ids_b}
df %>% 
  count(Vault_Scan_ID) %>% 
  count(n, name = "times_occuring")
```

So there is one `Vault_Scan_ID` that occurs twice, check which one:  
Vault_Scan_ID = 802688 occurs twice 

```{r, count_scan_ids_c}
df %>% 
  count(Vault_Scan_ID) %>% 
  filter(n > 1)
```

Look at the rows associated with that `Vault_Scan_ID` value along with the `Vault_UID` value.  

```{r, check_scan_ids_d}
vault_id_focus <- df %>% 
  count(Vault_Scan_ID) %>% 
  filter(n > 1) %>% 
  pull(Vault_Scan_ID)
df %>% 
  select(Vault_Scan_ID, Vault_UID) %>% 
  filter(Vault_Scan_ID %in% vault_id_focus)
```

Include columns involving date in the column name.  

```{r, check_scan_ids_e}
df %>% 
  select(Vault_Scan_ID, Vault_UID, ends_with("_date", ignore.case = TRUE)) %>% 
  filter(Vault_Scan_ID %in% vault_id_focus)
```

Check columns that contain "ID" in the column name.  

```{r, check_scan_ids_f}
df %>% 
  select(Vault_Scan_ID, Vault_UID, contains("ID", ignore.case = FALSE)) %>% 
  filter(Vault_Scan_ID %in% vault_id_focus)
```


**What other identifiers should be considered here??** 
Identifiers to include : age, sex, race, education , APOE_allele, PiB Regional values Left , PiB regional Right, PiB Global Values , GMD regional Left, GMD right, GMD global, FDG Regional left, FDG regional Right, FDG global  

How many unique people are there? And how many rows are associated with each unique person?  n = 93 ; max number of rows associated with a unique person is 4 

```{r, check_people_ids_a}
df %>% 
  count(Vault_UID) %>% 
  arrange(desc(n))
```

Count for the number of times each unique number of rows appears. So there are 6 `Vault_UID` values with 4 rows in the data set. There are 28 `Vault_ID` values with 3 rows in the data set. There are more unique `Vault_UID` values that appear more than once in the data set than the `Vault_UID` values that appear once.  

```{r, check_people_ids_b}
df %>% 
  count(Vault_UID) %>% 
  count(n, name='times_occurs')
```


## PiB Exploration

Extract the PiB related variables along with the unique identifiers.  

**Will need to do this again based on the unique identifiers that should be included.**  

```{r, get_pib_data}
pib_df <- df %>% select(Vault_Scan_ID, Vault_UID, starts_with("PIB_"))
pib_df %>% glimpse()
```

**Are the columns with names starting with `PiB_SUVR_GTM_` related to the "aggregate" or "macro" regions?**  

Yes, PiB_SUVR_GTM refers to the aggregate or global regions. 

For now, focus on the variables with `PiB_regional_` naming patterns.  

```{r, make_pib_regional_df}
pib_regional_wf <- pib_df %>% 
  select(starts_with("Vault"), PIB_Scan_Date, starts_with("PiB_regional"))
pib_regional_wf %>% glimpse()
```

Check number of missings per column.  

```{r, count_missings_per_column}
pib_regional_wf %>% purrr::map_dbl(~sum(is.na(.)))
```

Distribution of the PIB regional variables across all rows. This ignores any grouping by people.  

```{r, viz_pib_hists}
pib_regional_wf %>% 
  tibble::rowid_to_column() %>% 
  pivot_longer(cols = starts_with("PiB_regional")) %>% 
  ggplot(mapping = aes(x = value)) +
  geom_histogram(bins = 15) +
  facet_wrap(~name) +
  theme_bw() +
  theme(strip.background = element_blank(),
        strip.text = element_text(size = 4.5))
```


Use boxplots instead.  

```{r, viz_pib_boxplots}
pib_regional_wf %>% 
  tibble::rowid_to_column() %>% 
  pivot_longer(cols = starts_with("PiB_regional")) %>% 
  ggplot(mapping = aes(y = name, x = value)) +
  geom_boxplot(fill = 'grey') +
  theme_bw() +
  theme(axis.text.y = element_text(size = 5.5))
```


Start to parse the regional variable names.  

Create a helper function to separate the variable names based on the `'_'` pattern. The helper function is required because the number of underscores is not consistent across the regional names.  

```{r, make_helper_parser}
my_separate <- function(the_data, num_us)
{
  names_keep <- the_data %>% select(-name) %>% names()
  
  the_data %>% 
    tidyr::separate(name,
                    c("result_type", "regional_word",
                      sprintf("name_%02d", 1:(num_us-1))),
                    sep = "_") %>% 
    rename(!!!c("side_name" = sprintf("name_%02d", num_us-1))) %>% 
    tidyr::unite("region_name",
                 sprintf("name_%02d", 1:(num_us-2)),
                 sep = "_") %>% 
    select(all_of(c(names_keep,
                    "result_type", "region_name", "side_name")))
}
```

Apply the helper function based on the number of underscores to create the long-format data.  

```{r, check_parse_regional_names_a}
pib_regional_lf <- pib_regional_wf %>% 
  tibble::rowid_to_column() %>% 
  pivot_longer(cols = starts_with("PiB_regional")) %>% 
  mutate(num_underscores = stringr::str_count(name, pattern = "_")) %>% 
  group_by(num_underscores) %>% 
  tidyr::nest() %>% 
  mutate(ready_data = purrr::map2(data, num_underscores, my_separate)) %>% 
  select(num_underscores, ready_data) %>% 
  tidyr::unnest(ready_data) %>% 
  ungroup()
```


Check the unique region names and the number of rows associated with each region name.  

```{r, check_parse_regioinal_names_b}
pib_regional_lf %>% 
  ggplot(mapping = aes(y = region_name)) +
  geom_bar() +
  theme_bw()
```

Include the side.  

```{r, check_parse_regional_names_c}
pib_regional_lf %>% 
  ggplot(mapping = aes(y = region_name)) +
  geom_bar(mapping = aes(fill = side_name)) +
  scale_fill_brewer(palette = "Set1") +
  theme_bw() +
  theme(legend.position = 'top') 
```

Distributions of PIB values per region and side. Note the x-axis scales are the same in each facet.  

Frontal_Sup_2 regions shows significant asymmetry out of all the regions 

```{r, viz_pib_results_hist}
pib_regional_lf %>% 
  ggplot(mapping = aes(x = value)) +
  geom_freqpoly(mapping = aes(color = side_name,
                              group = interaction(region_name, 
                                                  side_name),
                              y = stat(density)),
                size = 1.2,
                bins = 15) +
  facet_wrap(~region_name, scales = "free_y") +
  scale_color_brewer("", palette = "Set1") +
  theme_bw() +
  theme(legend.position = 'top',
        axis.text.y = element_blank(),
        strip.background = element_blank(),
        strip.text = element_text(size = 6.5))
```

Allow the x-axis scales to vary across facets.  

```{r, viz_pib_results_hist_b}
pib_regional_lf %>% 
  ggplot(mapping = aes(x = value)) +
  geom_freqpoly(mapping = aes(color = side_name,
                              group = interaction(region_name, 
                                                  side_name),
                              y = stat(density)),
                size = 1.2,
                bins = 15) +
  facet_wrap(~region_name, scales = "free") +
  scale_color_brewer("", palette = "Set1") +
  theme_bw() +
  theme(legend.position = 'top',
        axis.text.y = element_blank(),
        strip.background = element_blank(),
        strip.text = element_text(size = 6.5))
```

Check the behavior of the repeated observations per person. The x-axis corresponds to a unique `Vault_UID` but the axis text has been turned off to make the plots easier to read.  

```{r, viz_result_per_person_a}
pib_regional_lf %>% 
  ggplot(mapping = aes(x = as.factor(Vault_UID), y = value)) +
  stat_summary(geom = 'linerange',
               fun.min = 'min', fun.max = 'max',
               mapping = aes(group = interaction(Vault_UID, 
                                                 region_name,
                                                 side_name),
                             color = side_name),
               position = position_dodge(0.1),
               size = 1.2, alpha = 0.5) +
  geom_point(mapping = aes(group = interaction(Vault_UID,
                                               region_name,
                                               side_name),
                           color = side_name),
             position = position_dodge(0.1)) +
  facet_wrap(~region_name, scales = 'free_y') +
  scale_color_brewer("", palette = "Set1") +
  theme_bw() +
  theme(legend.position = 'top',
        axis.text.x = element_blank(),
        strip.background = element_blank(),
        strip.text = element_text(size = 6.5))
```


Since the above figure is so busy, focus on a single region.  

Changed the region to Frontal_Sup_2 since it shows significant asymmetry 

```{r, viz_result_per_person_a_1}
pib_regional_lf %>% 
  filter(region_name %in% 'Frontal_Sup_2') %>% 
  ggplot(mapping = aes(x = as.factor(Vault_UID), y = value)) +
  stat_summary(geom = 'linerange',
               fun.min = 'min', fun.max = 'max',
               mapping = aes(group = interaction(Vault_UID, 
                                                 region_name,
                                                 side_name),
                             color = side_name),
               position = position_dodge(0.2),
               size = 1.2, alpha = 0.5) +
  geom_point(mapping = aes(group = interaction(Vault_UID,
                                               region_name,
                                               side_name),
                           color = side_name),
             position = position_dodge(0.2)) +
  facet_wrap(~region_name, scales = 'free_y') +
  scale_color_brewer("", palette = "Set1") +
  theme_bw() +
  theme(legend.position = 'top',
        axis.text.x = element_blank(),
        strip.background = element_blank(),
        strip.text = element_text(size = 13.5))
```

