title: "CS1950_Week05_GMD"
author: "Hunsi Jayaprakash"
date: "2/4/2022"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Load packages

```{r, load_tidyverse_pkg}
library(tidyverse)
```

```{r, read_csv_file}
file_name <- ("https://raw.githubusercontent.com/jyurko/cs_1950_brain_proj/main/KLU_APC2_Master_2021_07_07.csv?token=GHSAT0AAAAAABQQ7BMCJN2DI3DFUBFY6DSMYQXTN3A")
df <- readr::read_csv(file_name,col_names = TRUE)
```


## Check unique identifiers

The number of rows associated with each `Vault_Scan_ID`.  

```{r, count_scan_ids_a}
df %>% 
  count(Vault_Scan_ID)
```

The count for the number of rows associated with each `Vault_Scan_ID`.  

```{r, count_scan_ids_b}
df %>% 
  count(Vault_Scan_ID) %>% 
  count(n, name = "times_occuring")
```

So there is one `Vault_Scan_ID` that occurs twice, check which one:  
  Vault_Scan_ID = 802688 occurs twice 

```{r, count_scan_ids_c}
df %>% 
  count(Vault_Scan_ID) %>% 
  filter(n > 1)
```

Look at the rows associated with that `Vault_Scan_ID` value along with the `Vault_UID` value.  

```{r, check_scan_ids_d}
vault_id_focus <- df %>% 
  count(Vault_Scan_ID) %>% 
  filter(n > 1) %>% 
  pull(Vault_Scan_ID)
df %>% 
  select(Vault_Scan_ID, Vault_UID) %>% 
  filter(Vault_Scan_ID %in% vault_id_focus)
```


Include columns involving date in the column name.  

```{r, check_scan_ids_e}
df %>% 
  select(Vault_Scan_ID, Vault_UID, ends_with("_date", ignore.case = TRUE)) %>% 
  filter(Vault_Scan_ID %in% vault_id_focus)
```

Check columns that contain "ID" in the column name.  

```{r, check_scan_ids_f}
df %>% 
  select(Vault_Scan_ID, Vault_UID, contains("ID", ignore.case = FALSE)) %>% 
  filter(Vault_Scan_ID %in% vault_id_focus)
```

**What other identifiers should be considered here??** 
  Identifiers to include : age, sex, race, education , APOE_allele, PiB Regional values Left , PiB regional Right, PiB Global Values , GMD regional Left, GMD right, GMD global, FDG Regional left, FDG regional Right, FDG global  

How many unique people are there? And how many rows are associated with each unique person?  n = 93 ; max number of rows associated with a unique person is 4 

```{r, check_people_ids_a}
df %>% 
  count(Vault_UID) %>% 
  arrange(desc(n))
```

Count for the number of times each unique number of rows appears. So there are 6 `Vault_UID` values with 4 rows in the data set. There are 28 `Vault_ID` values with 3 rows in the data set. There are more unique `Vault_UID` values that appear more than once in the data set than the `Vault_UID` values that appear once.  

```{r, check_people_ids_b}
df %>% 
  count(Vault_UID) %>% 
  count(n, name='times_occurs')
```

## GMD Exploration

Extract the GMD related variables along with the unique identifiers.  

**Will need to do this again based on the unique identifiers that should be included.**  
  
```{r, get_GMD_data}
gmd_regional_df <- df %>% select(Vault_Scan_ID, Vault_UID, MR_Scan_Date, starts_with("GMD_"))
gmd_regional_df %>% glimpse()
```


```{r, get_gmd_global_data}
gmd_global_df <- df %>% select(Vault_Scan_ID, Vault_UID, starts_with("FS_"))
gmd_global_df %>% glimpse()
```


Focusing on the variables with `GMD_regional_` naming patterns.  

```{r, make_gmd_regional_df}
gmd_regional_wf <- gmd_regional_df %>% 
  select(starts_with("Vault"), MR_Scan_Date, starts_with("GMD_regional"))
gmd_regional_wf %>% glimpse()
```

Check number of missings per column.  

```{r, count_missings_per_column}
gmd_regional_wf %>% purrr::map_dbl(~sum(is.na(.)))
```

Distribution of the GMD regional variables across all rows. This ignores any grouping by people.  

```{r, viz_fdg_hists}
gmd_regional_wf %>% 
  tibble::rowid_to_column() %>% 
  pivot_longer(cols = starts_with("GMD_regional")) %>% 
  ggplot(mapping = aes(x = value)) +
  geom_histogram(bins = 10) +
  facet_wrap(~name) +
  theme_bw() +
  theme(strip.background = element_blank(),
        strip.text = element_text(size = 3.5))
```

Use boxplots instead.  

```{r, viz_GMD_boxplots}
gmd_regional_wf %>% 
  tibble::rowid_to_column() %>% 
  pivot_longer(cols = starts_with("GMD_regional")) %>% 
  ggplot(mapping = aes(y = name, x = value)) +
  geom_boxplot(fill = 'light blue') +
  theme_bw() +
  theme(axis.text.y = element_text(size = 5.5))
```

Start to parse the regional variable names.  

Create a helper function to separate the variable names based on the `'_'` pattern. The helper function is required because the number of underscores is not consistent across the regional names.  

```{r, make_helper_parser}
my_separate <- function(the_data, num_us)
{
  names_keep <- the_data %>% select(-name) %>% names()
  
  the_data %>% 
    tidyr::separate(name,
                    c("result_type", "regional_word",
                      sprintf("name_%02d", 1:(num_us-1))),
                    sep = "_") %>% 
    rename(!!!c("side_name" = sprintf("name_%02d", num_us-1))) %>% 
    tidyr::unite("region_name",
                 sprintf("name_%02d", 1:(num_us-2)),
                 sep = "_") %>% 
    select(all_of(c(names_keep,
                    "result_type", "region_name", "side_name")))
}
```

Apply the helper function based on the number of underscores to create the long-format data.  

```{r, check_parse_regional_names_a}
gmd_regional_lf <- gmd_regional_wf %>% 
  tibble::rowid_to_column() %>% 
  pivot_longer(cols = starts_with("GMD_regional")) %>% 
  mutate(num_underscores = stringr::str_count(name, pattern = "_")) %>% 
  group_by(num_underscores) %>% 
  tidyr::nest() %>% 
  mutate(ready_data = purrr::map2(data, num_underscores, my_separate)) %>% 
  select(num_underscores, ready_data) %>% 
  tidyr::unnest(ready_data) %>% 
  ungroup()
```


Check the unique region names and the number of rows associated with each region name.  

```{r, check_parse_regioinal_names_b}
gmd_regional_lf %>% 
  ggplot(mapping = aes(y = region_name)) +
  geom_bar() +
  theme_bw()
```

Include the side. 

Asymmetry is not significant in any of the regions 

```{r, check_parse_regional_names_c}
gmd_regional_lf %>% 
  ggplot(mapping = aes(y = region_name)) +
  geom_bar(mapping = aes(fill = side_name)) +
  scale_fill_brewer(palette = "Set1") +
  theme_bw() +
  theme(legend.position = 'top') 
```

Distributions of GMD values per region and side. Note the x-axis scales are the same in each facet.  

Some regions have larger left hemispheres in cortical volume. 

```{r, viz_GMD_results_hist}
gmd_regional_lf %>% 
  ggplot(mapping = aes(x = value)) +
  geom_freqpoly(mapping = aes(color = side_name,
                              group = interaction(region_name, 
                                                  side_name),
                              y = stat(density)),
                size = 1.2,
                bins = 15) +
  facet_wrap(~region_name, scales = "free_y") +
  scale_color_brewer("", palette = "Set1") +
  theme_bw() +
  theme(legend.position = 'top',
        axis.text.y = element_blank(),
        strip.background = element_blank(),
        strip.text = element_text(size = 6.5))
```

Allow the x-axis scales to vary across facets.
kept the bin number constant
zoom in to see the plot 

```{r, viz_GMD_results_hist_b}
gmd_regional_lf %>% 
  ggplot(mapping = aes(x = value)) +
  geom_freqpoly(mapping = aes(color = side_name,
                              group = interaction(region_name, 
                                                  side_name),
                              y = stat(density)),
                size = 1.2,
                bins = 15) +
  facet_wrap(~region_name, scales = "free") +
  scale_color_brewer("", palette = "Set1") +
  theme_bw() +
  theme(legend.position = 'top',
        axis.text.y = element_blank(),
        strip.background = element_blank(),
        strip.text = element_text(size = 3.5))
```

Check the behavior of the repeated observations per person. The x-axis corresponds to a unique `Vault_UID` but the axis text has been turned off to make the plots easier to read.  

```{r, viz_result_per_person_a}
gmd_regional_lf %>% 
  ggplot(mapping = aes(x = as.factor(Vault_UID), y = value)) +
  stat_summary(geom = 'linerange',
               fun.min = 'min', fun.max = 'max',
               mapping = aes(group = interaction(Vault_UID, 
                                                 region_name,
                                                 side_name),
                             color = side_name),
               position = position_dodge(0.1),
               size = 1.2, alpha = 0.5) +
  geom_point(mapping = aes(group = interaction(Vault_UID,
                                               region_name,
                                               side_name),
                           color = side_name),
             position = position_dodge(0.1)) +
  facet_wrap(~region_name, scales = 'free_y') +
  scale_color_brewer("", palette = "Set1") +
  theme_bw() +
  theme(legend.position = 'top',
        axis.text.x = element_blank(),
        strip.background = element_blank(),
        strip.text = element_text(size = 4.5))
```

Focused on a single region.  

Region: Frontal_Sup_2 since it showed significant FDG and PiB asymmetry. Did not show significant GMD asymmetry. 

```{r, viz_result_per_person_a_1}
gmd_regional_lf %>% 
  filter(region_name %in% 'Frontal_Sup_2') %>% 
  ggplot(mapping = aes(x = as.factor(Vault_UID), y = value)) +
  stat_summary(geom = 'linerange',
               fun.min = 'min', fun.max = 'max',
               mapping = aes(group = interaction(Vault_UID, 
                                                 region_name,
                                                 side_name),
                             color = side_name),
               position = position_dodge(0.2),
               size = 1.2, alpha = 0.5) +
  geom_point(mapping = aes(group = interaction(Vault_UID,
                                               region_name,
                                               side_name),
                           color = side_name),
             position = position_dodge(0.2)) +
  facet_wrap(~region_name, scales = 'free_y') +
  scale_color_brewer("", palette = "Set1") +
  theme_bw() +
  theme(legend.position = 'top',
        axis.text.x = element_blank(),
        strip.background = element_blank(),
        strip.text = element_text(size = 13.5))
```

Focused on a single region and one person.  

```{r, viz_result_per_person_a_2}
gmd_regional_lf %>% 
  filter(region_name %in% 'Frontal_Sup_2') %>% 
  ggplot(mapping = aes(x = as.factor(Vault_UID == 900720), y = value)) +
  stat_summary(geom = 'linerange',
               fun.min = 'min', fun.max = 'max',
               mapping = aes(group = interaction(Vault_UID, 
                                                 region_name,
                                                 side_name),
                             color = side_name),
               position = position_dodge(0.2),
               size = 1.2, alpha = 0.5) +
  geom_point(mapping = aes(group = interaction(Vault_UID,
                                               region_name,
                                               side_name),
                           color = side_name),
             position = position_dodge(0.2)) +
  facet_wrap(~region_name, scales = 'free_y') +
  scale_color_brewer("", palette = "Set1") +
  theme_bw() +
  theme(legend.position = 'top',
        axis.text.x = element_blank(),
        strip.background = element_blank(),
        strip.text = element_text(size = 13.5))
```

**Do you want to focus on a single person?**  
  
  The filter should be applied before the ggplot() call.  

```{r, viz_result_per_person_a_2_bb}
gmd_regional_lf %>% 
  filter(region_name %in% 'Frontal_Sup_2') %>% 
  filter(Vault_UID %in% c(900720)) %>% 
  ggplot(mapping = aes(x = as.factor(Vault_UID), y = value)) +
  stat_summary(geom = 'linerange',
               fun.min = 'min', fun.max = 'max',
               mapping = aes(group = interaction(Vault_UID, 
                                                 region_name,
                                                 side_name),
                             color = side_name),
               position = position_dodge(0.2),
               size = 1.2, alpha = 0.5) +
  geom_point(mapping = aes(group = interaction(Vault_UID,
                                               region_name,
                                               side_name),
                           color = side_name),
             position = position_dodge(0.2)) +
  facet_wrap(~region_name, scales = 'free_y') +
  scale_color_brewer("", palette = "Set1") +
  theme_bw() +
  theme(legend.position = 'top',
        axis.text.x = element_blank(),
        strip.background = element_blank(),
        strip.text = element_text(size = 13.5))
```

```{r, viz_result_per_person_a_3}
gmd_spread <- gmd_regional_df%>%
  select(starts_with("GMD_regional_Frontal_Sup_2"))
gmd_spread%>% summary()
gmd_regional_lf %>% bind_rows() %>% 
  group_by(Vault_UID) %>% summarise(num_rows = n(), avg_x = mean(value), .groups = 'drop') %>% summary(avg_x)
```

## Vault_UID ranges

Calculate the range of the value for each person in each region per side. Show for a single region and person. First isolate that person and region.  

```{r, check_range_calc_a1}
gmd_regional_lf %>% 
  filter(region_name %in% 'Frontal_Sup_2') %>% 
  filter(Vault_UID %in% c(900720)) %>% 
  select(Vault_Scan_ID, Vault_UID, MR_Scan_Date, result_type, region_name, side_name, value)
```


Then group and summarize.  

```{r, check_range_calc_a2}
gmd_regional_lf %>% 
  filter(region_name %in% 'Frontal_Sup_2') %>% 
  filter(Vault_UID %in% c(900720)) %>% 
  group_by(result_type, region_name, side_name, Vault_UID) %>% 
  summarise(num_rows = n(),
            num_missing = sum(is.na(value)),
            num_scans = n_distinct(Vault_Scan_ID),
            first_scan = min(MR_Scan_Date),
            last_scan = max(MR_Scan_Date),
            avg_value = mean(value, na.rm = TRUE),
            min_value = min(value, na.rm = TRUE),
            max_value = max(value, na.rm = TRUE),
            .groups = 'drop')
```


Calculate the range and show the summary stats for this one person and region.  

```{r, check_range_calc_a3}
gmd_regional_lf %>% 
  filter(region_name %in% 'Frontal_Sup_2') %>% 
  filter(Vault_UID %in% c(900720)) %>% 
  group_by(result_type, region_name, side_name, Vault_UID) %>% 
  summarise(num_rows = n(),
            num_missing = sum(is.na(value)),
            num_scans = n_distinct(Vault_Scan_ID),
            first_scan = min(MR_Scan_Date),
            last_scan = max(MR_Scan_Date),
            avg_value = mean(value, na.rm = TRUE),
            min_value = min(value, na.rm = TRUE),
            max_value = max(value, na.rm = TRUE),
            .groups = 'drop') %>% 
  mutate(range_value = max_value - min_value) %>% 
  select(region_name, side_name, Vault_UID, num_scans, num_missing, ends_with("_value"))
```


Calculate the summary stats for all regions and people.  

The below code chunk generates a ton of warnings!  
  
```{r, make_range_calc_a1}
gmd_regional_summary <- gmd_regional_lf %>% 
  group_by(result_type, region_name, side_name, Vault_UID) %>% 
  summarise(num_rows = n(),
            num_missing = sum(is.na(value)),
            num_scans = n_distinct(Vault_Scan_ID),
            first_scan = min(MR_Scan_Date),
            last_scan = max(MR_Scan_Date),
            avg_value = mean(value, na.rm = TRUE),
            min_value = min(value, na.rm = TRUE),
            max_value = max(value, na.rm = TRUE),
            .groups = 'drop') %>% 
  mutate(range_value = max_value - min_value)
```


Why the warnings? Check the missings.  

```{r, check_range_calc_missings_a1}
gmd_regional_summary %>% 
  mutate(num_non_missing = num_rows - num_missing) %>% 
  filter(num_non_missing == 0) %>% 
  select(num_rows, num_missing, num_scans, num_non_missing, min_value, max_value)
```

How many people have missing value issues?  8 people with missing values 
  
```{r, check_range_calc_missings_a2}
gmd_regional_summary %>% 
  mutate(num_non_missing = num_rows - num_missing) %>% 
  filter(num_non_missing == 0) %>% 
  group_by(Vault_UID) %>% 
  summarise(num_regions = n_distinct(region_name),
            .groups = 'drop')
```

Look at one of those people.  

```{r, check_range_calc_missings_a3}
gmd_regional_summary %>% 
  filter(Vault_UID == 901469) %>% 
  select(Vault_UID, region_name, side_name, num_rows, num_missing, ends_with("_value"))
```

Look at the summary stats for the distribution of the range across people in each region and side. Remove the people that have the missings issue. 

Frontal regions have a bigger range. As well the R Precentral region 

```{r, viz_range_results_a1}
gmd_regional_summary %>% 
  filter(!is.infinite(range_value)) %>% 
  ggplot(mapping = aes(y = region_name, x = range_value)) +
  geom_boxplot(mapping = aes(group = interaction(region_name, side_name))) +
  facet_wrap(~side_name) +
  theme_bw()
```

rank the people based on the range in the Frontal_Sup_2 region.  

```{r, rank_people_one_region_range_a1}
people_ranking_order <- gmd_regional_summary %>% 
  filter(!is.infinite(range_value)) %>% 
  filter(region_name == "Frontal_Sup_2", side_name == "R") %>% 
  arrange(desc(range_value)) %>% 
  pull(Vault_UID)
```


Look at the top 5 people based on the range in the one region. 901883 has a larger range. 

```{r}
gmd_regional_lf %>% 
  filter(region_name %in% 'Frontal_Sup_2') %>% 
  filter(Vault_UID %in% people_ranking_order[1:5]) %>% 
  ggplot(mapping = aes(x = as.factor(Vault_UID), y = value)) +
  stat_summary(geom = 'linerange',
               fun.min = 'min', fun.max = 'max',
               mapping = aes(group = interaction(Vault_UID, 
                                                 region_name,
                                                 side_name),
                             color = side_name),
               position = position_dodge(0.2),
               size = 1.2, alpha = 0.5) +
  geom_point(mapping = aes(group = interaction(Vault_UID,
                                               region_name,
                                               side_name),
                           color = side_name),
             position = position_dodge(0.2)) +
  facet_wrap(~region_name, scales = 'free_y') +
  scale_color_brewer("", palette = "Set1") +
  theme_bw() +
  theme(legend.position = 'top',
        strip.background = element_blank(),
        strip.text = element_text(size = 13.5))
```


Include the average and standard error on the average for each person.  

```{r}
gmd_regional_lf %>% 
  filter(region_name %in% 'Frontal_Sup_2') %>% 
  filter(Vault_UID %in% people_ranking_order[1:5]) %>% 
  ggplot(mapping = aes(x = as.factor(Vault_UID), y = value)) +
  stat_summary(geom = 'linerange',
               fun.min = 'min', fun.max = 'max',
               mapping = aes(group = interaction(Vault_UID, 
                                                 region_name,
                                                 side_name),
                             color = side_name),
               position = position_dodge(0.2),
               size = 1.7, alpha = 0.5) +
  geom_point(mapping = aes(group = interaction(Vault_UID,
                                               region_name,
                                               side_name),
                           color = side_name),
             position = position_dodge(0.2),
             size = 3) +
  stat_summary(fun.data = 'mean_se',
               mapping = aes(group = interaction(Vault_UID,
                                                 region_name,
                                                 side_name)),
               fun.args = list(mult = 2),
               color = 'black',
               position = position_dodge(0.2)) +
  facet_wrap(~region_name, scales = 'free_y') +
  scale_color_brewer("", palette = "Set1") +
  theme_bw() +
  theme(legend.position = 'top',
        strip.background = element_blank(),
        strip.text = element_text(size = 13.5))
```


Include the people with the 5 lowest range as well for comparison purposes.  

```{r}
gmd_regional_lf %>% 
  filter(region_name %in% 'Frontal_Sup_2') %>% 
  filter(Vault_UID %in% c(people_ranking_order[1:5],
                          rev(people_ranking_order)[1:5])) %>% 
  ggplot(mapping = aes(x = as.factor(Vault_UID), y = value)) +
  stat_summary(geom = 'linerange',
               fun.min = 'min', fun.max = 'max',
               mapping = aes(group = interaction(Vault_UID, 
                                                 region_name,
                                                 side_name),
                             color = side_name),
               position = position_dodge(0.2),
               size = 1.7, alpha = 0.5) +
  geom_point(mapping = aes(group = interaction(Vault_UID,
                                               region_name,
                                               side_name),
                           color = side_name),
             position = position_dodge(0.2),
             size = 3) +
  stat_summary(fun.data = 'mean_se',
               mapping = aes(group = interaction(Vault_UID,
                                                 region_name,
                                                 side_name)),
               fun.args = list(mult = 2),
               color = 'black',
               position = position_dodge(0.2)) +
  facet_wrap(~region_name, scales = 'free_y') +
  scale_color_brewer("", palette = "Set1") +
  theme_bw() +
  theme(legend.position = 'top',
        strip.background = element_blank(),
        strip.text = element_text(size = 13.5))
```
